Parsing Postfix log files

### 2009-01-13 doc/logparser-other-parsers-review.tex
Further corrections and improvements after proofreading.

Each of the reviewed parsers was tested with the \numberOFlogFILES{} test
log files described in \sectionref{parser efficiency}.  The data extracted
by \parsername{} is documented in \sectionref{connections table} and
\sectionref{results table}; for convenience that list is repeated here:
server \gls{IP}, server hostname, client \gls{IP}, client hostname, helo
hostname, queueid, start time, end time, \gls{SMTP} code, sender,
recipient, size, message ID\@.

@@

Whether a MySQL database or Berkeley DB table is chosen in addition to the
\gls{CSV} output, \gls{LMA} stores the following data: time and date,
client hostname and \gls{IP} address, server \gls{IP} address, sender and
recipient addresses, \gls{SMTP} code, and size (for accepted mails only).
It is unclear which time and date is stored: start time, end time, or
delivery time?  Unlike \parsername{} it does not store the server hostname,
helo hostname, queueid, start and end times, timestamps for each log line,
or message id (for accepted mails only).  Handling of multiple recipients,
\gls{SMTP} codes, or remote servers\footnote{A single mail may be sent to
multiple remote servers if it was addressed to recipients in different
domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained; experimental observation shows that multiple
records are added when there are multiple recipients (sadly the records are
not associated or linked in any way), and presumably the same approach is
taken when there are multiple destination servers.

@@

Pflogsumm produces a report designed for troubleshooting rather than
in-depth analysis.  It does not support saving any data, and it does not
extract any data than is not required for producing its report.  Both the
parsing and reporting are difficult to extend as it is a specialised tool,
unlike the easily extensible design of \parsername{}.  It does not
correlate log lines by queueid or \gls{pid}, and does not need to deal with
the complications encountered during this project.  Pflogsumm produces a
useful report, and successfully parsed the \numberOFlogFILES{} log files
tested with,\footnote{The results it reported were not verified in detail,
but it did not report any errors, and has a very good reputation amongst
Postfix users.} but it is not a suitable base for this project.  It does
not extract the helo hostname, queueid, start and end times, timestamps for
each log line, or message id.

Pflogsumm has many options are available to include or exclude certain
sections of the report; by default it includes the following:

@@

Sawmill is a general purpose commercial product that parses 818 log file
formats (as of 2008/11/23) and produces reports from the extracted data.
Its data extraction facilities (described later) are too limited to save
sufficient data for the purposes of this project: although it can extract
three different sets of data from Postfix log files, they are not
interlinked in any way.  The documentation does not suggest that any
attempt is made to correlate log lines or deal with the difficulties
documented in \sectionref{complications}.

@@

    \item \url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{}
        Fields extracted: from, to, server, UID, relay, status, number of
        recipients, origin hostname, origin \gls{IP}, and virus.  It also
        counts the number of and total size of all mails delivered.  The
        fields \texttt{server}, \texttt{uid}, \texttt{relay}, and
        \texttt{virus} are not explained in the documentation:
        \texttt{server} is probably the server the mail is delivered to,
        and \texttt{uid} might be the uid of the user submitting mail
        locally.  Postfix does not perform any form of virus checking
        (though it has many options for cooperating with an external virus
        scanner), so the \texttt{virus} field is a mystery.

    \item \url{http://www.thesawmill.co.uk/formats/postfix_ii.html}
        \newline{} Fields extracted: from, to, RBL list, client hostname,
        and client \gls{IP}\@.  It also counts the number of and total size
        of all mails delivered.

@@

Even if the three data sets were combined Sawmill would extract less data
than \parsername{}: it omits the helo hostname, queueid, and start and end
times.  Sawmill does not extract any data about rejections except when the
rejection is caused by a \gls{DNSBL} check (\texttt{RBL list} in the list
of fields).

@@

Sawmill's web interface supports searching on any combination of the fields
it extracts, and the results obtained from searches were accurate.  The
interface for searching is neither as simple to use nor as informative as
the interface provided by Splunk.  The administrative interface is much
easier to use than Splunk's: it took only five minutes to start parsing a
whole directory of log files.  Searches performed using the web interface
produced reasonable results, which were not verified in detail.

@@

In summary Sawmill suffers from being a general purpose product; it is
probably much more useful when parsing log files where each log line is
self-contained, rather than linked to other log lines (e.g.\ web server log
files).  It is not suitable as a base for this parser, as the source code
made available is obfuscated and not intended for modification; in addition
the architecture would probably need to be overhauled to deal with
correlating log lines.

@@

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface supports complicated searches, providing statistics and
graphs in real time, a facility not provided by \parsername{} (report
generation has been deferred to a subsequent program).  Saved searches can
be run periodically and the results emailed to a recipient or sent to an
external program for further processing (though maybe without the graphs
and detailed statistics); the author was unable to save searches, though
that may have been due to limitations in the free version.  The database is
not available for use by external programs, whereas \parsername{} provides
the database and leaves it to the user to utilise it without limit or
restriction.  The interface is optimised for interactive use rather than
automated queries and it does not appear to be possible to write
independent tools to utilise the Splunk database; there are additional
reports available on \url{http://www.splunkbase.com/}, but the author was
unable to find any: every category was empty, even those that the interface
claimed had reports available.  Many types of reports are bundled with the
software, though most are variations of a bar or pie chart, except bubble
and heatmap graphs.  It is easy to drill down through the graphs to extract
a portion of the data (e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific address),
though it is not possible to search on partial words.  All searches
performed using the indexed data returned reasonable results.  The web
interface is quite attractive and simple to use when searching, but as an
administrator it seems unnecessarily difficult to perform simple tasks.

When testing Splunk it took roughly 30 minutes to figure out how to add a
single log file to be indexed so that it could be searched, with the
downside that the log file was copied into a spool directory before
indexing, doubling the disk space usage.  The next test was to index all
the log files in a particular directory, but after three hours, numerous
futile attempts, and reading all the available documentation, the author
admitted defeat.  Using the \gls{CLI} rather than the web interface was
more successful: invoking the command \newline{} \tab{} \texttt{splunk find
logs }\textit{log-directory\/}\newline{} added 40 of the
\numberOFlogFILES{} log files to the queue for indexing.  Further attempts
enqueued the same 40 log files, without explaining why the others were
excluded.\footnote{The log files appear to have been indexed once only;
presumably Splunk keeps track of the files it has indexed and discards
requests to index files for a second time.  This may or may not be a useful
feature for \parsername{}.} There did not appear to be an option to ensure
the log files would be processed in the order they were created, though
this may be neither necessary nor beneficial with Splunk.  Subsequently the
author was successful in adding a single file at a time using the
\gls{CLI}; a simple loop to add all desired log files was sufficient to
index all files.  Splunk will periodically check all those files for
updates unless they are manually removed from its list; this may or may not
be useful behaviour.  Splunk did not appear to have any difficulty in
indexing the log files, once they had been successfully added to its queue.
\parsername{} parses the logs it is instructed to parse, in the order
given; periodic parsing of logs is a task an administrator can easily
achieve with \texttt{cron(8)} and \texttt{logrotate(8)}.

Copious documentation is made available on \url{http://www.splunk.com/},
but poor organisation and sheer abundance makes it extremely hard to find
useful information.  Searching using the website's interface confusingly
tends to return results from old documentation rather than new.  In general
the documentation appears to have been written by someone intimately
acquainted with the software who has difficulty understanding how a
newcomer would approach tasks or the questions they would ask.

@@

When parsing Postfix log files Splunk extracts some standard fields: to and
from addresses, HELO hostname, time and date, the host the log files were
generated on, and protocol (\gls{SMTP} or \gls{ESMTP}).  It parses the
standard syslog\glsadd{syslog} fields at the beginning of the log line, and
extracts any \texttt{key=<value>} pairs occurring after the standard syslog
prologue; these pairs are the fields listed above.  Searches can be based
on the extracted fields or the full text of the log line.  \parsername{}
extracts noticeably more data (client and server \gls{IP} and hostname,
queueid, start and end times, timestamps for each log line, \gls{SMTP}
code, and message ID), though it does not make the full text of the line
available (this could be trivially added if desired, but would greatly
increase the size of the resulting database).  The full power of \gls{SQL}
is available when searching the data extracted by \parsername{}, allowing
the user to search on arbitrarily complicated conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=<value>} fields from a log line; most
importantly it makes no attempt to correlate log lines by queueid or
\gls{pid}, or to handle any of the myriad complications discussed in
\sectionref{complications}.  Some additional Postfix reports are supposedly
available at \url{http://www.splunkbase.com/}, but the author was unable to
find Postfix reports, or additional reports for any other log file types.

@@

Isoqlog's report misses most of the information gathered by \parsername{},
and it only reports on the domains listed in its configuration file, making
it impossible to produce reports for every sender domain.  It ignores all
log lines except those for today, so it is impossible to analyse historical
logs, and testing with the \numberOFlogFILES{} test log files is pointless.
It does maintain a record of data previously extracted, which the newly
extracted data is merged into (no information is provided on the format of
the data store).  The data extracted is limited to the number of mails sent
by each sender, unlike the copious amounts of data extracted by
\parsername{}.  It does not utilise rejection log lines in any way, so is
unsuitable for the purposes of this project.  Its parsing is completely
inextensible, indeed is almost incomprehensible, relying on
\texttt{scanf(3)}, unexplained fixed offsets, and low level string
manipulation; it is the opposite end of the spectrum to \parsernames{}
parsing.  It does not handle any of the complications discussed in
\sectionref{complications}, does not gather the breadth of data required
for this project, and ignores the majority of log lines produced by
Postfix.

@@

\noindent{}YAALA's Postfix parser extracts some of the fields \parsername{}
does: it stores either the \gls{IP} or the hostname for client and server,
not both; it omits the helo hostname, queueid, \gls{SMTP} code, size of
each accepted mail, start and end times, timestamps for each log line, and
message ID\@.  It extracts some data that \parsername{} does not: the delay
in delivering each mail, and how many times delivery was deferred for each
mail; these could easily be extracted by \parsername{} if desired.  XXX
EXTRACT DELAY AND DEFER ONCE AUTOMATIC EXTRACTION HAS BEEN FINISHED\@.
Unlike \parsername{}, YAALA does not maintain separate counters for each
restriction; this rules out the possibility of using the collected data for
optimisation, testing or understanding of restrictions.  YAALA's Postfix
parser does not deal with the complications explained in
\sectionref{complications}, though it does correlate log lines by queueid.

@@

\noindent{}Examining the source code reveals that the parser looks for
\texttt{<key>=<value>} pairs in each log line, extracts them, and
correlates the data by queueid.  This approach will find the following
fields: helo hostname, queueid, \gls{SMTP} code, sender and recipient
addresses, and size of accepted mails.  It is unclear if the parser will
extract any further data.  Lire misses the following fields extracted by
\parsername{}: client and server \gls{IP} and hostname, start and end
times, timestamps of each log line, and message ID\@.

@@

It also counts the number of log lines parsed and skipped.  It omits client
\gls{IP} and hostname, server \gls{IP}, helo hostname, queueid, and message
ID\@.  It extracts the delay for delivered mails, which \parsername{} does
not.  Log lines are correlated based on the queueid (referred to as
sessionname [sic] within Logrep), but not by \gls{pid}.  The parsing is
error prone: empty fields are saved when the log line does not match the
regex, though it appears that they will not overwrite existing data.  Most
notably rejections are completely ignored, making it unsuitable for the
purposes of this project.  It does not try to address any of the
complications in \sectionref{complications} except for correlating by
queueid.

>>>>

Each of the reviewed parsers was tested with the \numberOFlogFILES{} test
log files described in \sectionref{parser efficiency}.  The data extracted
by \parsername{} is documented in \sectionref{connections table} and
\sectionref{results table}; for convenience that list is repeated here:
server \gls{IP}, server hostname, client \gls{IP}, client hostname, HELO
hostname, queueid, start time, end time, \gls{SMTP} code, sender,
recipient, size, message ID\@.

@@

Whether a MySQL database or Berkeley DB table is chosen in addition to the
\gls{CSV} output, \gls{LMA} stores the following data: time and date,
client hostname and \gls{IP} address, server \gls{IP} address, sender and
recipient addresses, \gls{SMTP} code, and size (for accepted mails only).
It is unclear which time and date is stored: start time, end time, or
delivery time?  Unlike \parsername{} it does not store the server hostname,
HELO hostname, queueid, start and end times, timestamps for each log line,
or message id (for accepted mails only).  Handling of multiple recipients,
\gls{SMTP} codes, or remote servers\footnote{A single mail may be sent to
multiple remote servers if it was addressed to recipients in different
domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained; experimental observation shows that multiple
records are added when there are multiple recipients (sadly the records are
not associated or linked in any way), and presumably the same approach is
taken when there are multiple destination servers.

@@

Pflogsumm produces a report designed for troubleshooting rather than
in-depth analysis.  It does not support saving any data, and it does not
extract any data that it does not require to produce its report, e.g.\ it
does not extract the HELO hostname, queueid, start and end times,
timestamps for each log line, or message id.  Both the parsing and
reporting are difficult to extend because it is a specialised tool, unlike
the easily extensible design of \parsername{}.  It does not correlate log
lines by queueid or \gls{pid}, and does not need to deal with the
complications encountered during this project (\sectionref{complications}).
Pflogsumm produces a useful report, and successfully parsed the
\numberOFlogFILES{} log files it was tested with.\footnote{The results it reported
were not verified in detail, but it did not report any errors, and has a
very good reputation amongst Postfix users.}
Pflogsumm has many options to include or exclude certain
sections of the report; by default it includes the following:

@@

Sawmill is a general purpose commercial product that parses 818 log file
formats (as of 2008/11/23) and produces reports from the extracted data.
Its data extraction facilities (described later) are too limited to save
sufficient data for the purposes of this project: although it can extract
three different sets of data from Postfix log files, they are not
interlinked in any way.  The documentation does not suggest that any
attempt is made to correlate log lines by either queueid or pid or to deal
with the difficulties documented in \sectionref{complications}.

@@

    \item \url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{}
        Fields extracted: from, to, server, UID, relay, status, number of
        recipients, origin hostname, origin \gls{IP}, and virus.  It also
        counts the number of and total size of all mails delivered.  The
        fields \texttt{server}, \texttt{uid}, \texttt{relay}, and
        \texttt{virus} are not explained in the documentation:
        \texttt{server} is probably the hostname or \gls{IP} of the server
        the mail is delivered to; \texttt{relay} might be the delivery
        method: \gls{SMTP}, local delivery, or \gls{LMTP}; \texttt{uid}
        might be the uid of the user submitting mail locally.  Postfix does
        not perform any form of virus checking (though it has many options
        for cooperating with an external virus scanner), so the
        \texttt{virus} field is a mystery.

    \item \url{http://www.thesawmill.co.uk/formats/postfix_ii.html}
        \newline{} Fields extracted: from, to, RBL list, client hostname,
        and client \gls{IP}\@.  It also counts the number and total size of
        all mails delivered.

@@

Even if the three data sets were combined Sawmill would extract less data
than \parsername{}: it omits the HELO hostname, queueid, and start and end
times.  Sawmill does not extract any data about rejections except when the
rejection is caused by a \gls{DNSBL} check (\texttt{RBL list} in the list
of fields).

@@

Sawmill's web interface supports searching on any combination of the fields
it extracts, and searches performed using the web interface produced
accurate results.  The interface for searching is neither as simple to use
nor as informative as the interface provided by Splunk.  The administrative
interface is much easier to use than Splunk's: it took only five minutes to
start parsing a whole directory of log files.

@@

In summary Sawmill suffers from being a general purpose product; it is
probably much more useful when parsing log files where each log line is
self-contained (e.g.\ web server log files), rather than log files
containing interlinked log lines.  It is not suitable as a base for this
parser, as the source code made available is obfuscated and not intended
for modification; in addition the architecture would probably need to be
overhauled or replaced to deal with correlating log lines.

@@

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface supports complicated searches, providing statistics and
graphs in real time, a facility not provided by \parsername{} (report
generation has been deferred to a subsequent program).  Searches can be
based on the fields extracted by Splunk or the full text of the log line.
Splunk allows quite complicated searches, but does not make the raw data
available in an accessible form.  Saved searches can be run periodically
and the results emailed to a recipient or sent to an external program for
further processing (though maybe without the graphs and detailed
statistics); the author was unable to save searches, though that may have
been due to limitations in the free version.  The database is not available
for use by external programs, whereas \parsername{} provides the database
and leaves it to the user to utilise it without limit or restriction.  The
interface is optimised for interactive use rather than automated queries
and it does not appear to be possible to write independent tools to utilise
the Splunk database;.  Some additional Postfix reports are supposedly
available at \url{http://www.splunkbase.com/}, but the author was unable to
find any Postfix reports, or indeed for any other log file types: every
category was empty, even those that the web site claimed had many reports
available.  Many types of graphs can be generated, though most are
variations of a bar or pie chart, except bubble and heatmap graphs.  It is
easy to drill down through the graphs to extract a portion of the data
(e.g.\ select the hour with the largest number of events, then select a
particular host, and finally a specific address), though it is not possible
to search on partial words.  All searches performed using the indexed data
returned reasonable results.

The web interface is quite attractive and simple to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing Splunk it took roughly 30 minutes to figure out how to
add a single log file to be indexed so that it could be searched, with the
downside that the log file was copied into a spool directory before
indexing, doubling the disk space usage.  The next test was to index all
the log files in a particular directory, but after three hours, numerous
futile attempts, and reading all the available documentation, the author
admitted defeat.  Using the \gls{CLI} rather than the web interface was
partially successful: the command \newline{} \tab{} \texttt{splunk find
logs }\textit{log-directory\/}\newline{} added 40 of the
\numberOFlogFILES{} log files to the queue for indexing.  Further attempts
enqueued the same 40 log files, without explaining why the others were
excluded.\footnote{The log files appear to have been indexed once only;
presumably Splunk keeps track of the files it has indexed and discards
requests to index files for a second time.  This may or may not be a useful
feature for \parsername{}.} There did not appear to be an option to ensure
the log files would be processed in the order they were created, though
this may be neither necessary nor beneficial with Splunk.  Subsequently the
author was successful in adding a single file at a time using the
\gls{CLI}:\newline{} \tab{} \texttt{splunk add tail
}\textit{filename\/}\newline{} A simple loop was then enough to add all the
desired log files.  Splunk will periodically check all indexed files for
updates unless they are manually removed from its list; this may or may not
be useful behaviour.  Splunk did not appear to have any difficulty in
indexing the log files, once they had been successfully added to its queue.
\parsername{} parses the logs it is instructed to parse, in the order
given; periodic parsing of logs is a task an administrator can easily
achieve with \texttt{cron(8)} and \texttt{logrotate(8)}.

Copious documentation is made available on \url{http://www.splunk.com/},
but poor organisation and sheer abundance makes it extremely hard to find
useful information.  Searches confusingly tended to return results from old
documentation rather than new.  In general the documentation appears to
have been written by someone intimately acquainted with the software, who
has difficulty understanding how a newcomer would approach tasks or the
questions they would ask.

@@

When parsing Postfix log files Splunk parses the standard
syslog\glsadd{syslog} fields at the beginning of the log line, and extracts
any \texttt{key=value} pairs occurring after the standard syslog prologue:
to and from addresses, HELO hostname, and protocol (\gls{SMTP} or
\gls{ESMTP}).  \parsername{} extracts noticeably more data (client and
server \gls{IP} and hostname, queueid, start and end times, timestamps for
each log line, \gls{SMTP} code, and message ID), though it does not make
the full text of the line available (this could be trivially added if
desired, but would greatly increase the size of the resulting database).
The full power of \gls{SQL} is available when searching the data extracted
by \parsername{}, allowing the user to search on arbitrarily complicated
conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=value} fields from a log line; most
importantly it makes no attempt to correlate log lines by queueid or
\gls{pid}, or to handle any of the other myriad complications discussed in
\sectionref{complications}.  Its source code is unavailable, so it could
not be used as a base for this project, even if it fulfilled all other
requirements.

@@

Isoqlog's report misses most of the information gathered by \parsername{}:
the data extracted is limited to the number of mails sent by each sender,
and it only reports on senders from the domains listed in its configuration
file, making it impossible to produce complete reports.  It ignores all log
lines except those with today's date, so it is impossible to analyse
historical log files, and testing with the \numberOFlogFILES{} test log
files was pointless.  It does maintain a record of data previously
extracted, which the newly extracted data is merged into; the format of the
data store is undocumented.  It does not utilise rejection log lines in any
way, so is unsuitable for the purposes of this project.  Its parsing is
completely inextensible, indeed is almost incomprehensible, relying on
\texttt{scanf(3)}, unexplained fixed offsets, and low level string
manipulation; it is the opposite end of the spectrum to \parsernames{}
parsing.  It does not handle any of the complications discussed in
\sectionref{complications}, does not gather the breadth of data required
for this project, and ignores the majority of log lines produced by
Postfix.

@@

\noindent{}YAALA's Postfix parser extracts some of the fields \parsername{}
does: it stores either the \gls{IP} or the hostname for client and server,
not both; it omits the HELO hostname, queueid, \gls{SMTP} code, size of
each accepted mail, start and end times, timestamps for each log line, and
message ID\@.  It extracts some data that \parsername{} does not: the delay
in delivering each mail, and how many times delivery was deferred for each
mail; these could easily be extracted by \parsername{} if desired.  XXX
EXTRACT DELAY AND DEFER ONCE AUTOMATIC EXTRACTION HAS BEEN FINISHED\@.
Unlike \parsername{}, YAALA does not maintain separate counters for each
restriction; this rules out the possibility of using the collected data for
optimisation, testing or understanding of restrictions.  YAALA's Postfix
parser does not deal with the complications explained in
\sectionref{complications}, though it does correlate log lines by queueid.

@@

\noindent{}Examining the source code reveals that the parser looks for
\texttt{<key>=<value>} pairs in each log line, extracts them, and
correlates the data by queueid.  This approach will find the following
fields: HELO hostname, queueid, \gls{SMTP} code, sender and recipient
addresses, and size of accepted mails.  It is unclear if the parser will
extract any further data.  Lire misses the following fields extracted by
\parsername{}: client and server \gls{IP} and hostname, start and end
times, timestamps of each log line, and message ID\@.

@@

It also counts the number of log lines parsed and skipped.  It omits client
\gls{IP} and hostname, server \gls{IP}, HELO hostname, queueid, and message
ID\@.  It extracts the delay for delivered mails, which \parsername{} does
not.  Log lines are correlated based on the queueid (referred to as
sessionname [sic] within Logrep), but not by \gls{pid}.  The parsing is
error prone: empty fields are saved when the log line does not match the
regex, though it appears that they will not overwrite existing data.  Most
notably rejections are completely ignored, making it unsuitable for the
purposes of this project.  It does not try to address any of the
complications in \sectionref{complications} except for correlating by
queueid.
